### 2023-08-31 sections/2-methods.tex
changes to rnn

[TODO: delete this and talk about overlapping children improvement instead]
We conduct $\rho$-nearest neighbors search as described in CHESS~\cite{ishaq2019clustered}, but 
with some implementation improvements including identification of candidate clusters which are leaves and 
conducting exhaustive search over those leaves.

\begin{algorithm} 
    \caption{$\rho$-NN(\emph{clusters, query, r})} 
    \label{alg:rnn} 
    \begin{algorithmic}[2]
        \REQUIRE $r \geq 0$
        \REQUIRE $clusters \neq \emptyset$
        \STATE $results \leftarrow \emptyset$
        \IF{$clusters.left$}
            \IF{$clusters.left.\delta$ $\leq$ $r$ + $clusters.left.radius$}
                \STATE $\rho$-NN($clusters.left, query, r$)
            \ENDIF
        \ENDIF
        \IF{$clusters.right$}
            \IF{$clusters.right.\delta$ $\leq$ $r$ + $clusters.right.radius$} 
                \STATE $\rho$-NN($clusters.right, query, r$)
            \ENDIF
        \ENDIF
        \IF{$\neg$$clusters.left$ $\land \neg$$clusters.right$}
            \FOR{$p \in clusters.points$}
                \IF{$p.\delta \leq r$}
                    \STATE $results \leftarrow r$
                \ENDIF
            \ENDFOR
        \ENDIF
        \STATE Return $results$
    \end{algorithmic}
    \end{algorithm}

The asymptotic complexity of $\rho$-nearest neighbors is the same as in CHESS~\cite{ishaq2019clustered}.

@@

In these algorithms, we typically use $H$, for \emph{hits}, to refer to the data structure which stores the closest points to the query found so far.

@@

Once $|H| > 0$, we continue to perform $\rho$-NN search, but instead of 
increasing the radius by a factor of 2 on each iteration, we increase it by a factor determined 
by the local fractal dimension of the clusters containing the points found so far. In particular, 
we increase the radius by a factor of 
\begin{equation} \text{argmin}\left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right) \label{2} \end{equation}
where $\mu$ is the local fractal dimension (abbreviated lfd in algorithm below) in the region around the query point.
Intuitively, the factor by which we increase the radius should be \emph{inversely} related to the number of points found so far. 
Additionally, when the local fractal dimension at the radius scale from the previous iteration is low, this suggests that the data 
are relatively concentrated in that region; thus, the factor of radius increase should be \emph{directly} related to the 
local fractal dimension. Once $|H| >= k$, we return the $k$ closest points.

\begin{algorithm} % enter the algorithm environment
    \caption{Repeated$\rho$-NN(\emph{tree, query, k})} % give the algorithm a caption
    \label{alg:knn-by-rnn} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[4] % enter the algorithmic environment
        \STATE $H \leftarrow$ $\emptyset$
        \STATE $m \leftarrow$ $\frac{root.radius}{root.cardinality}$
        \WHILE {$|H| = 0$}
            \STATE $H.push(\rho$-NN$(tree, query, m)$)
            \STATE $m \leftarrow 2m$
        \ENDWHILE
        \WHILE {$|H| < k$}
            \STATE $clusters \leftarrow \{ C: \exists p \in H \land p \in C \}$
            \STATE $\mu \leftarrow \frac{1}{|clusters|} \sum_{C \in clusters} C.lfd$
            \STATE $m \leftarrow \text{min}\left(2, \frac{k}{|H|^{\frac{1}{\mu}}}\right)$
            \STATE $H.push(\rho$-NN$(tree, query, m)$)
        \ENDWHILE
        \STATE $H.sort()$
        \STATE Return $H[0], H[1], \cdots H[k-1]$
    \end{algorithmic}
    \end{algorithm}

>>>>

We conduct $\rho$-nearest neighbors search as described in \cite{ishaq2019clustered}, but 
with the following improvement: when a cluster overlaps with the query ball, instead of  
always proceeding with both of its children, we only proceed with those children which 
also overlap with the query ball.

\begin{algorithm} 
    \caption{$\rho$-NN(\emph{clusters, query, r})} 
    \label{alg:rnn} 
    \begin{algorithmic}[2]
        \REQUIRE $r \geq 0$
        \REQUIRE $clusters \neq \emptyset$
        \STATE $results \leftarrow \emptyset$
        \IF{$clusters.left \land clusters.left \cap B_X(query, r) \neq \emptyset$}
            \IF{$clusters.left.\delta$ $\leq$ $r$ + $clusters.left.radius$}
                \STATE $\rho$-NN($clusters.left, query, r$)
            \ENDIF
        \ENDIF
        \IF{$clusters.right \land clusters.right \cap B_X(query, r) \neq \emptyset$}
            \IF{$clusters.right.\delta$ $\leq$ $r$ + $clusters.right.radius$} 
                \STATE $\rho$-NN($clusters.right, query, r$)
            \ENDIF
        \ENDIF
        \IF{$\neg$$clusters.left$ $\land \neg$$clusters.right$}
            \FOR{$p \in clusters.points$}
                \IF{$p.\delta \leq r$}
                    \STATE $results.push((p, r))$
                \ENDIF
            \ENDFOR
        \ENDIF
        \STATE Return $results$
    \end{algorithmic}
    \end{algorithm}

The asymptotic complexity of $\rho$-nearest neighbors is the same as in ~\cite{ishaq2019clustered}.

@@

In these algorithms, we typically use $H$, for \emph{hits}, to refer to the data structure which stores the closest points to the query found so far and
$Q$ to refer to the data structure which stores the clusters and points which are still in contention for being one of the nearest neighbors.

@@

Once $|H| > 0$, we continue to perform $\rho$-NN search, but instead of 
increasing the radius by a factor of 2 on each iteration, we increase it by a factor determined 
by the local fractal dimension of the clusters containing the points found so far. In particular, 
we increase the radius by a factor of 
\begin{equation} \text{arg min}\left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right) \label{2} \end{equation}
where $\mu$ is the local fractal dimension (abbreviated lfd in algorithm below) in the region around the query point.
Intuitively, the factor by which we increase the radius should be \emph{inversely} related to the number of points found so far. 
Additionally, when the local fractal dimension at the radius scale from the previous iteration is low, this suggests that the data 
are relatively concentrated in that region; thus, the factor of radius increase should be \emph{directly} related to the 
local fractal dimension. Once $|H| >= k$, we return the $k$ closest points.

\begin{algorithm} % enter the algorithm environment
    \caption{Repeated$\rho$-NN(\emph{tree, query, k})} % give the algorithm a caption
    \label{alg:knn-by-rnn} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[4] % enter the algorithmic environment
        \STATE $H \Leftarrow$ $\emptyset$
        \STATE $m \Leftarrow$ $\frac{root.radius}{root.cardinality}$
        \WHILE {$|H| = 0$}
            \STATE $H.push(\rho$-NN$(tree, query, m)$)
            \STATE $m \Leftarrow 2m$
        \ENDWHILE
        \WHILE {$|H| < k$}
            \STATE $Q \Leftarrow \{ C: \exists p \in H \land p \in C \}$
            \STATE $\mu \leftarrow \frac{1}{|Q|} \sum_{C \in Q} C.lfd$
            \STATE $m \leftarrow \text{arg min}\left(2, \left({\frac{k}{|H|}}\right)^{\frac{1}{\mu}}\right)$
            \STATE $H.push(\rho$-NN$(tree, query, m)$)
        \ENDWHILE
        \STATE $H.sort()$
        \STATE Return $H[0], H[1], \cdots H[k-1]$
    \end{algorithmic}
    \end{algorithm}
